---
title: "Data manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

# Load in the FAS litters data 

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

# Let's take a look at select 
Allows us to look at some columns and not others 

```{r}
select(litters_df,group,gd0_weight)
```

# You can also specify a range of columns

```{r}
select(litters_df,group,gd0_weight:gd_of_birth)
```

# You can get rid of a column, this gets rid of litter number
```{r}
select(litters_df,litter_number)
```

Renaming colums
```{r}
select(litters_df,GROUP = group, LITTER_NUMBER = litter_number)
```

You can simply rename without using the select function
```{r}
rename(litters_df,GROUP = group, LITTER_NUMBER = litter_number)
```

Select helpers
```{r}
select(litters_df,starts_with("gd"))
```
You can move things and then keep order for everything else
```{r}
select(litters_df, litter_number,everything())
```
relocate does something similar
```{r}
relocate(litters_df, group)
```

## Filter: intended to keep rows in which you are interested in keeping for subsequent analysis; sekect operates on columns, filter operates on rows
```{r}
filter(litters_df,gd0_weight >= 22)
```

You can ask if gd_of_birth ==20, you use more than one equal sign because you are not trying to say that the variable is = the number 20. It is a test for equality
```{r}
filter(litters_df,gd_of_birth == 20)
```

Not equal to 20, you can use exclamation marks
```{r}
filter(litters_df,!gd_of_birth == 20)
```

```{r}
filter(litters_df, gd0_weight >= 22,gd_of_birth == 20)
```

```{r}
filter(litters_df,group == "Mod8")
```

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## Mutate: allows you to modify or create additiional variables
```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```

## Arrange 
```{r}
arrange(litters_df,pups_born_alive)
```

## %>%
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, pups_survive)
litters_mutated = mutate (litters_data_selected,wt_gain = gd18_weight-gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0weight)
```

USE THE PIPES OPERATOR INSTEAD
```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>%  
  janitor::clean_names()
  select(-pups_survive)
  mutate(wt_gain = gd18_weight - gd0_weight)
  drop_na(gd0_weight)
```

