---
title: "Data manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

# Load in the FAS litters data 

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

# Let's take a look at select 
Allows us to look at some columns and not others 

```{r}
select(litters_df,group,gd0_weight)
```

# You can also specify a range of columns

```{r}
select(litters_df,group,gd0_weight:gd_of_birth)
```

# You can get rid of a column, this gets rid of litter number
```{r}
select(litters_df,litter_number)
```

Renaming colums
```{r}
select(litters_df,GROUP = group, LITTER_NUMBER = litter_number)
```

You can simply rename without using the select function
```{r}
rename(litters_df,GROUP = group, LITTER_NUMBER = litter_number)
```

Select helpers
```{r}
select(litters_df,starts_with("gd"))
```
You can move things and then keep order for everything else
```{r}
select(litters_df, litter_number,everything())
```
relocate does something similar
```{r}
relocate(litters_df, group)
```

